Paper Report

### 2017-11-16 Paper_report.tex
Save

\item Purpose: Find $\theta$ which maximize p(X $|\theta$)
\item How: Use Generalized EM to estimate $\theta$
	\begin{enumerate}
	\item : Maximizing $p(X)$ means
	\begin{equation}
	p(x) = \sum_{z} p(X,Z) = E_{q(Z)}[\ln p(X,Z|\theta)] + H[q(Z)] + KL(q(Z) || p(Z|X,\theta))
	\end{equation}
	\item E-step: Estimate $p(Z|X)$ based on current $\theta$
	\item M-step: Consider $q(Z) = p(Z|X)$ \\
	$KL(q(Z) || p(Z|X,\theta)) = 0$ \\
	$H[q(Z)]$ constant
	\item Maximizing $p(X)$ becomes maximizing $E_{q(Z)}[\ln p(X,Z|\theta)]$
	\end{enumerate}
\item Result: Gradient of the joint probability scaled by the posteriors
	\begin{equation}
	J(\theta) = \sum_{Z} p(Z|X) \dfrac{\partial}{\partial\theta} \ln p(X,Z| \theta)
	\label{Tran16unsupervisedGradientJointProbability} 
	\end{equation}
\item Problem: How to calculate $p(X,Z)$ ?
\end{itemize}

@@

\item Therefore, the probability of a given sequence of observation X and latent variables Z (Factorization of the joint probability):
	\begin{equation}
	p(X,Z) = \prod_{t=1}^{n + 1} p(z_{t}|z_{t-1})\prod_{t=1}^{n} p(x_{t}|z_{t})
	\label{Tran16unsupervisedProbabilityHMM}
	\end{equation}

\item Result: Combine $p(X,Z)$ \eqref{Tran16unsupervisedProbabilityHMM} and gradient $J(\theta)$ \eqref{Tran16unsupervisedGradientJointProbability}

@@

@@

\begin{equation}
s_{NN}(a|e, f) = \prod_{i = 1}^{|e|} t_{lex}(e_{i},f_{a_{i}}|e,f) t_{d}(a_{i},a_{i-1}|e,f)
\label{Yang13wordNNAlignmentScore}
\end{equation}

	\begin{itemize}
	\item $s_{NN}(a|e, f)$ is the score of an alignment based on the source and target sentence.
	\item $f_{a_{i}}$ represents not only this word but the context around this word. Surrounding words of both source and target word are input of DNN, which is handling the context of both sides. This could reduce the explosion of parameter number. \\ In this case, they use fixed length windows surrounding both $e_{i}$ and $f_{j}$. 
	\item $t_{lex}$: Lexical translation score.
	\begin{equation}
	t_{lex}(e_{i},f_{j}|e,f) = functions_{NN}(window(e_{i}),window(f_{j}) )
	\label{Yang13wordLexicalTranslationScore}
	\end{equation}
	\item $t_{d}$: Distortion score.
	\begin{equation}
	t_{d}(a_{i}, a_{i-1}|e,f) = t_{d}(a_{i} - a_{i-1}|window(f_{a_{i-1}})) = functions_{NN}(window(f_{a_{i-1}}))
	\end{equation}
	They recognize that this lexicalized distortion does not produce a better alignment. They reverse to the simple version.
	\begin{equation}
	t_{d}(a_{i}, a_{i-1}|e,f) = t_{d}(a_{i} - a_{i-1})
	\end{equation}
	\end{itemize}

\subsection{Loss in supervised learning}
\begin{equation}
loss(\theta) = \sum_{every(\textbf{e},\textbf{f})} \max(0 , 1 - s_{\theta}(a^+ |\textbf{e},\textbf{f}) + s_{\theta}(a^- |\textbf{e},\textbf{f}) )
\label{Yang13wordAlignmentLoss}
\end{equation}
where
\begin{itemize}
\item $a^+$: Gold alignment path
\item $a^-$: Highest scoring incorrect alignment path under $\theta$
\item $s_\theta$: Score for alignment path defined in \eqref{Yang13wordNNAlignmentScore}
\end{itemize}

\subsection{Analysis}
\subsubsection{Two loss functions during training}
They pre-train word embedding with monolingual data.
They train firstly NN with the loss from lexical translation score and then the loss \eqref{Yang13wordAlignmentLoss} because they recognize that using only the loss \eqref{Yang13wordAlignmentLoss} is not efficient.

The loss of lexical translation score:
\begin{equation}
loss(\theta) = \sum_{every(\textbf{e},\textbf{f})} \max(0 , 1 - t_{lex, \theta}((e,f)^+ | \textbf{e},\textbf{f}) + t_{lex, \theta}((e,f)^- | \textbf{e},\textbf{f}) )
\end{equation}
where
\begin{itemize}
\item $(e,f)^+$: Correct word pair
\item $(e,f)^-$: Incorrect word pair
\item $t_{lex}$: Score of \eqref{Yang13wordLexicalTranslationScore}
\end{itemize}

@@

\subsection{Loss in unsupervised learning} 
\begin{equation}
loss(\theta) = max(0 , 1 - \sum_{(\textbf{f}^+,\textbf{e}^+) \in T} E_{\Phi} [s_{\theta} (\textbf{a}|\textbf{f}^+, \textbf{e}^+)]  + \sum_{(\textbf{f}^+,\textbf{e}^-) \in \Omega} E_{\Phi} [s_{\theta} (\textbf{a}|\textbf{f}^+, \textbf{e}^-)] )
\label{Tamura14recurrentLossUnsupervisedLearning}
\end{equation}
where
\begin{itemize}
\item T: Training data (Bilingual sentences) as Observed data
\item $\Omega$ : Full translation search space as neighbourhood of observe data, .
\item $\Phi$ : Set of all possible alignments given (\textbf{e},\textbf{f})
\item $E_{\Phi} [s_{\theta}]$ : Expected value of the scores $s_\theta$ on $\Phi$
\item $e^+$ : A target language sentence in T
\item $e^-$ : A pseudo-target language sentence
\item $\sum_{(\textbf{f}^+,\textbf{e}^+) \in T} E_{\Phi} [s_{\theta} (\textbf{a}|\textbf{f}^+, \textbf{e}^+)]$ : Expectation term for observed data
\item$ \sum_{(\textbf{f}^+,\textbf{e}^-) \in \Omega} E_{\Phi} [s_{\theta} (\textbf{a}|\textbf{f}^+, \textbf{e}^-)] )$ : Expectation term for neighbourhood
\end{itemize}

They would like to reduce the computation, they randomly select N pseudo-target language sentences for each $f^+$. \eqref{Tamura14recurrentLossUnsupervisedLearning} becomes:
\begin{equation}
loss(\theta) = \sum_{\textbf{f}^+ \in T} max(0 , 1 -  E_{GEN} [s_{\theta} (\textbf{a}|\textbf{f}^+, \textbf{e}^+)]  + \frac{1}{N} \sum_{\textbf{e}^-} E_{GEN} [s_{\theta} (\textbf{a}|\textbf{f}^+, \textbf{e}^-)] )
\end{equation}
where
\begin{itemize}
\item $\textbf{e}^-$ : a pseudo-target language sentence with the same length $|\textbf{e}^-| = |\textbf{e}^+|$
\item GEN: A subset of all possible alignment generated by beam search
\end{itemize}

\subsection{Agreement constraints} 
\subsection{Analysis}

@@

\subsection{Hidden Markov Alignment Model}
The alignment model is re-structured:
\begin{equation}
p(f_1^J, a_1^J|e_1^I) = p(J|e_1^I) \prod_{j=1}^J p(f_j, a_j | f_1^{j-1}, a_1^{j-1}, e_1^I)
\end{equation}
\begin{equation}
= p(J|e_1^I) \prod_{j=1}^J p( f_j | f_1^{j-1}, a_j, e_1^I) * p( a_j | f_1^{j-1}, a_1^{j-1}, e_1^I)
\end{equation}

@@

\begin{equation}
p(f_1^J, a_1^J|e_1^I) = p(J|I) \prod_{j=1}^J p(f_j | e_{a_j}) * p(a_j | a_{j-1}, I)
\label{Och2003SystematicAlignmentModelEquation}
\end{equation}

@@

\section{Word alignment model in Neural Network}
We start from this formula \eqref{Och2003SystematicAlignmentModelEquation}:

\begin{equation}
p(f_1^J, a_1^J|e_1^I) = p(J|I) \prod_{j=1}^J p(f_j | e_{a_j}) * p(a_j | a_{j-1}, I)
\end{equation}

In general, we try to calculate these probability by using two different neural networks.
The significant point of NN is Context.

\subsection{Probability approach}
The papers include \cite{Tran16unsupervised}.

\subsection{Non-probability - Score approach}
The papers include \cite{Yang13word}, \cite{Tamura14recurrent}.

>>>>

\item Purpose: Find $\theta$ which maximize p(X $|\theta$)
\item How: Use Generalized EM to estimate $\theta$
	\begin{enumerate}
	\item : Maximizing $p(X)$ means
	\begin{equation}
	p(x) = \sum_{z} p(X,Z) = E_{q(Z)}[\ln p(X,Z|\theta)] + H[q(Z)] + KL(q(Z) || p(Z|X,\theta))
	\end{equation}
	\item E-step: Estimate $p(Z|X)$ based on current $\theta$
	\item M-step: Consider $q(Z) = p(Z|X)$ \\
	$KL(q(Z) || p(Z|X,\theta)) = 0$ \\
	$H[q(Z)]$ constant
	\item Maximizing $p(X)$ becomes maximizing $E_{q(Z)}[\ln p(X,Z|\theta)]$
	\end{enumerate}
\item Result: Gradient of the joint probability scaled by the posteriors
	\begin{equation}
	J(\theta) = \sum_{Z} p(Z|X) \dfrac{\partial}{\partial\theta} \ln p(X,Z| \theta)
	\label{Tran16unsupervised Gradient Joint Probability} 
	\end{equation}
\item Problem: How to calculate $p(X,Z)$ ?
\end{itemize}

@@

\item Therefore, the probability of a given sequence of observation X and latent variables Z (Factorization of the joint probability):
	\begin{equation}
	p(X,Z) = \prod_{t=1}^{n + 1} p(z_{t}|z_{t-1})\prod_{t=1}^{n} p(x_{t}|z_{t})
	\label{Tran16unsupervised Probability HMM}
	\end{equation}

\item Result: Combine $p(X,Z)$ \eqref{Tran16unsupervised Probability HMM} and gradient $J(\theta)$ \eqref{Tran16unsupervised Gradient Joint Probability}

@@

%----------------------------------------------------------------------------------------
%	Alignment by agreement
%----------------------------------------------------------------------------------------
\section{Alignment by Agreement \citep{Liang2006Alignment}}

@@

\begin{equation}
s_{NN}(a|e, f) = \prod_{i = 1}^{|e|} t_{lex}(e_{i},f_{a_{i}}|e,f) t_{d}(a_{i},a_{i-1}|e,f)
\label{Yang13word NN Alignment Score}
\end{equation}

	\begin{itemize}
	\item $s_{NN}(a|e, f)$ is the score of an alignment based on the source and target sentence.
	\item $f_{a_{i}}$ represents not only this word but the context around this word. Surrounding words of both source and target word are input of DNN, which is handling the context of both sides. This could reduce the explosion of parameter number. \\ In this case, they use fixed length windows surrounding both $e_{i}$ and $f_{j}$. 
	\item $t_{lex}$: Lexical translation score.
	\begin{equation}
	t_{lex}(e_{i},f_{j}|e,f) = functions_{NN}(window(e_{i}),window(f_{j}) )
	\label{Yang13word Lexical Translation Score}
	\end{equation}
	\item $t_{d}$: Distortion score.
	\begin{equation}
	t_{d}(a_{i}, a_{i-1}|e,f) = t_{d}(a_{i} - a_{i-1}|window(f_{a_{i-1}})) = functions_{NN}(window(f_{a_{i-1}}))
	\end{equation}
	They recognize that this lexicalized distortion does not produce a better alignment. They reverse to the simple version.
	\begin{equation}
	t_{d}(a_{i}, a_{i-1}|e,f) = t_{d}(a_{i} - a_{i-1})
	\end{equation}
	\end{itemize}

\subsection{Loss in supervised learning}
\begin{equation}
loss(\theta) = \sum_{every(\textbf{e},\textbf{f})} \max(0 , 1 - s_{\theta}(a^+ |\textbf{e},\textbf{f}) + s_{\theta}(a^- |\textbf{e},\textbf{f}) )
\label{Yang13word Alignment Loss}
\end{equation}
where
\begin{itemize}
\item $a^+$: Gold alignment path
\item $a^-$: Highest scoring incorrect alignment path under $\theta$
\item $s_\theta$: Score for alignment path defined in \eqref{Yang13word NN Alignment Score}
\end{itemize}

\subsection{Analysis}
\subsubsection{Two loss functions during training}
They pre-train word embedding with monolingual data.
They train firstly NN with the loss from lexical translation score and then the loss \eqref{Yang13word Alignment Loss} because they recognize that using only the loss \eqref{Yang13word Alignment Loss} is not efficient.

The loss of lexical translation score:
\begin{equation}
loss(\theta) = \sum_{every(\textbf{e},\textbf{f})} \max(0 , 1 - t_{lex, \theta}((e,f)^+ | \textbf{e},\textbf{f}) + t_{lex, \theta}((e,f)^- | \textbf{e},\textbf{f}) )
\end{equation}
where
\begin{itemize}
\item $(e,f)^+$: Correct word pair
\item $(e,f)^-$: Incorrect word pair
\item $t_{lex}$: Score of \eqref{Yang13word Lexical Translation Score}
\end{itemize}

@@

\subsection{Loss in unsupervised learning} 
\begin{equation}
loss(\theta) = max(0 , 1 - \sum_{(\textbf{f}^+,\textbf{e}^+) \in T} E_{\Phi} [s_{\theta} (\textbf{a}|\textbf{f}^+, \textbf{e}^+)]  + \sum_{(\textbf{f}^+,\textbf{e}^-) \in \Omega} E_{\Phi} [s_{\theta} (\textbf{a}|\textbf{f}^+, \textbf{e}^-)] )
\label{Tamura14recurrent Loss Unsupervised Learning}
\end{equation}
where
\begin{itemize}
\item T: Training data (Bilingual sentences) as Observed data
\item $\Omega$ : Full translation search space as neighbourhood of observe data, .
\item $\Phi$ : Set of all possible alignments given (\textbf{e},\textbf{f})
\item $E_{\Phi} [s_{\theta}]$ : Expected value of the scores $s_\theta$ on $\Phi$
\item $e^+$ : A target language sentence in T
\item $e^-$ : A pseudo-target language sentence
\item $\sum_{(\textbf{f}^+,\textbf{e}^+) \in T} E_{\Phi} [s_{\theta} (\textbf{a}|\textbf{f}^+, \textbf{e}^+)]$ : Expectation term for observed data
\item$ \sum_{(\textbf{f}^+,\textbf{e}^-) \in \Omega} E_{\Phi} [s_{\theta} (\textbf{a}|\textbf{f}^+, \textbf{e}^-)] )$ : Expectation term for neighbourhood
\end{itemize}

They would like to reduce the computation, they randomly select N pseudo-target language sentences for each $f^+$. \eqref{Tamura14recurrent Loss Unsupervised Learning} becomes:
\begin{equation}
loss(\theta) = \sum_{\textbf{f}^+ \in T} max(0 , 1 -  E_{GEN} [s_{\theta} (\textbf{a}|\textbf{f}^+, \textbf{e}^+)]  + \frac{1}{N} \sum_{\textbf{e}^-} E_{GEN} [s_{\theta} (\textbf{a}|\textbf{f}^+, \textbf{e}^-)] )
\end{equation}
where
\begin{itemize}
\item $\textbf{e}^-$ : a pseudo-target language sentence with the same length $|\textbf{e}^-| = |\textbf{e}^+|$
\item GEN: A subset of all possible alignment generated by beam search
\end{itemize}

\subsection{Agreement constraints}
\marginpar{See \citep{Liang2006Alignment}}

Why: HMM-based model is asymmetric. It is demonstrated that encouraging directional models to agree improves alignment performance.
\\
How: They propose training two directional models including alignment F$\rightarrow$E and E$\rightarrow$F. Each model has a different modified loss function

\begin{equation}
loss_{\theta_{F \rightarrow E}} = loss(\theta_{F \rightarrow E})  + \alpha \parallel \theta_{E \rightarrow F} - \theta_{F \rightarrow E} \parallel
\end{equation}

\begin{equation}
loss_{\theta_{E \rightarrow F}} = loss(\theta_{E \rightarrow F})  + \alpha \parallel \theta_{F \rightarrow E} - \theta_{E \rightarrow F} \parallel
\end{equation}

where
\begin{itemize}
\item $\theta_{E \rightarrow F}$ : Weight of layers in model $E \rightarrow F$
\item $\alpha$: Weight parameter of the agreement constraint $\parallel \theta_{E \rightarrow F} - \theta_{F \rightarrow E} \parallel$
\item $\parallel \theta \parallel$: Norm of $\theta$
\item $loss(\theta_{F \rightarrow E})$: Loss from \eqref{Tamura14recurrent Loss Unsupervised Learning}
\end{itemize}

\subsection{Analysis} 
Baselines are IBM4, FFNN \citep{Yang13word}. They do the experiments of word alignment and machine translation in both unsupervised/supervised learning.

\subsubsection{Result}
RNNs captures alignment paths based on long alignment history. This can be viewed as phrase-level alignment, which is more effective in non-similar-order-language (Japanese-English) than French-English.\\

RNNs could help to reduce the size of training data. They show the similar results between RNNs-Small data and IBM4-Large data.\\

They highlight the role of agreement constraints in alignment performance improvement.\\

@@

\subsection{Hidden Markov Alignment Model \cite{Och2003Systematic}}
The alignment model is re-structured:
\begin{equation}
p(f_1^J, a_1^J|e_1^I) = p(J|e_1^I) \prod_{j=1}^J p(f_j, a_j | f_1^{j-1}, a_1^{j-1}, e_1^I)
\end{equation}
\begin{equation}
= p(J|e_1^I) \prod_{j=1}^J p( f_j | f_1^{j-1}, a_j, e_1^I) * p( a_j | f_1^{j-1}, a_1^{j-1}, e_1^I)
\end{equation}

@@

\begin{equation}
p(f_1^J, a_1^J|e_1^I) = p(J|I) \prod_{j=1}^J p(f_j | e_{a_j}) * p(a_j | a_{j-1}, I)
\label{Och2003Systematic Alignment Model Equation}
\end{equation}

@@

\section{Word alignment model in Neural Network}
We start from this formula \eqref{Och2003Systematic Alignment Model Equation}:

\begin{equation}
p(f_1^J, a_1^J|e_1^I) = p(J|I) \prod_{j=1}^J p(f_j | e_{a_j}) * p(a_j | a_{j-1}, I)
\label{Word Alignment Model Neural Network HMM Equation}
\end{equation}

In general, we try to calculate these probability by using two different neural networks.
The significant point of NN is embedding context.

\subsection{Probability approach}
Unsupervised neural HMM of \cite{Tran16unsupervised} shows how to apply NN in HMM. \eqref{Word Alignment Model Neural Network HMM Equation} becomes:
\begin{equation}
p(f_1^J, a_1^J|e_1^I) = \prod_{j=1}^J p(f_j | e_{a_j}) * p(a_j | a_{j-1}, I)
\end{equation}

\subsection{Non-probability - Score approach}
The papers are \cite{Yang13word}, \cite{Tamura14recurrent}.

\section{Techniques in Alignment}

\subsection{Agreement between models}
